---
title: "Class09 Work"
author: "Justin"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Analyzing Data**

```{r}
# SECTION 1
# Importing Data (Wisconsin Cancer Data)

url <- "https://bioboot.github.io/bimm143_W18/class-material/WisconsinCancer.csv"

wisc.df <- read.csv(url)

# Convert features to matrix (ONLY TAKE IMPORTANT COLUMNS)

wisc.data <- as.matrix(wisc.df[,3:32])

head(wisc.data)

```


There are `r nrow(wisc.df)` patients in this study.


**Tabulating samples**
Benign (B) vs Malignant (M) tumor samples.

```{r}
table(wisc.df$diagnosis)
```


```{r}

# Set the row names of wisc.data

row.names(wisc.data) <- wisc.df$id

# Store diagnosis separately

diagnosis <- wisc.df$diagnosis

```


## Exploring the Data First

Q1. How many observations are in this dataset?
Q2. How many variables/features in the data are suffixed with _mean?
Q3. How many of the observations have a malignant diagnosis?


```{r}
# 1; rows = observations since each row is a patient

nrow(wisc.data)

# 2; Find PATTERN with grep(pattern,x)

colnames(wisc.df)

x <- grep("_mean", colnames(wisc.df))

# Use length to get how many of the columns end w _mean

length(x)


# 3; use table() to count those of each col

table(wisc.df$diagnosis)


```

## Section 2
### **Performing PCA**
```{r}

# SECTION 2

# Check column means and standard deviations; make it look pretty with round()
colMeans(wisc.data)

apply(wisc.data,2,sd)

round(colMeans(wisc.data), 3)

round(apply(wisc.data,2,sd), 3)




```

Values look hella different so we want to use `scale = TRUE` when we do PCA.


```{r}

# Do PCA on wisc.data

wisc.pr <- prcomp(wisc.data, scale = TRUE)



```



Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

**44.27%**

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

**3; need to go to PC3**

Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

**7; need to go to PC7**


```{r}
# Look at summary of results; this'll help answer the questions from above

summary(wisc.pr)

```


## Interpret PCA Results
**GOTTA PLOT IT**

Plot using `plot()`; gonna need to call the matrix `wisc.pr` and call `x`; then specify column 1 for PC1 and column 2 for PC2. 

Then, color the plot by the `diagnosis` using `col = diagnosis`.

```{r}


plot(wisc.pr$x[,1], wisc.pr$x[,2],col=diagnosis, pch=1, xlab = "PC1", ylab = "PC2")


plot(wisc.pr$x[, c(1, 3)], col = (diagnosis), 
     xlab = "PC1", ylab = "PC3")
```



**UGLY PLOT -> Nice plot**
```{r}
# Plot

biplot(wisc.pr)

# PC1 v PC3

plot(wisc.pr$x[,1], wisc.pr$x[,3], col=diagnosis, pch=1, xlab = "PC1", ylab = "PC3")

```


Do a different figure

```{r}

wisc.matrix <- dist(wisc.pr$x[,1:4])

wisc.pr.hclust <- hclust(d=wisc.matrix, method = "ward.D2")

wisc.pr.hclust

plot(wisc.pr.hclust, main = "Dendrogram of Patients" )

grps <- cutree(wisc.pr.hclust, k = 2)

# Table to see how much is in each whatever thingy

table(grps)

table(grps, diagnosis)



```




**SKIPPED HELLA**
..........


## Section 7

```{r}
# SECTION 7
# Import Data

url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata = new)

npc

# Plot the stuff

plot(wisc.pr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")

# Red = malignant; Black = benign BC RED BAD


```




