---
title: "Class 14"
author: "Justin"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Importing `countData` and `colData` 

```{r}

# Import dataaa

counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)

# Look @ data

head(counts)

head(metadata)



```


Comparing the averages of the control data v averages of the experimental data to determine if there's actually a difference between the two.

```{r}
# call upon specifically the control

control <- metadata[metadata[,"dex"]=="control",]


control$id


```


Now calculae mean values across control columns **countdata**

```{r}

counts[,control$id]

# get sums

control.mean <- rowSums(counts[,control$id])/length(control$id)

names(control.mean) <- counts$ensgene

```



Do it for the other data

```{r}

treated <- metadata[metadata[,"dex"]=="treated",]


treated$id

counts[,treated$id]

# get sums

treated.mean <- rowSums(counts[,treated$id])/length(treated$id)

names(treated.mean) <- counts$ensgene



```


Combine meancount data for bookkeeping purposes

```{r}

meancounts <- data.frame(control.mean, treated.mean)

# colSums to better compare data instead of raw data

colSums(meancounts)

```

Plot control v treated

```{r}

# Plotting the control v treated to see the relationship between them

plot(meancounts$control.mean, meancounts$treated.mean)



```


Plot in log scale to better see the relationship;
add log = "xy" to plot function to plot in log scale

```{r}

# log plot

meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)


plot(meancounts$control.mean, meancounts$treated.mean, log = "xy")


```


Remove the rows that have weird results: NaN, -Inf, 0, etc; because we can't say anything about this data either

```{r}
# Testing out the arr.ind thing

x <- data.frame(apples=c(69, 15, 22, 69), pears=c(17, 12, 69, 69))

x

which(x==69, arr.ind = TRUE)


```



```{r}
# The arr.ind tells us at what location there's a value that we are looking for; in this case, it will tell us what row and what col there's a value that = 0, and we're assigning all of the things to zero.vals

zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)

```




A common threshold used for calling something differentially expressed is a log2(FoldChange) of greater than 2 or less than -2. Let’s filter the dataset both ways to see how many genes are up or down-regulated.

```{r}

up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)


# get a count of how many are above(up) or below(down) 
sum(up.ind)

sum(down.ind)

```


## Using DESeq2 Analysis

```{r, eval = FALSE}

library(DESeq2)
citation("DESeq2")

```

We will use the `DESeqDataSetFromMatrix()` function to build the required DESeqDataSet object and call it dds, short for our DESeqDataSet. If you get a warning about “some variables in design formula are characters, converting to factors” don’t worry about it. Take a look at the dds object once you create it

```{r}

# Tells DESeq2 to go thru all of the stuff 

dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds


# Dunno what these do

sizeFactors(dds)

dispersions(dds)

# results(dds)

# 

dds <- DESeq(dds)

# Getting results

res <- results(dds)
res

summary(res)


```



Volcano Plot

```{r}

# 

res$sig <- res$padj<0.05 & abs(res$log2FoldChange)>2

# How many of each?
table(res$sig)


```

```{r}
# Tellsus how many NAs there are in res$sig

sum(is.na(res$sig))

```



Plotting

```{r}
# Set the color palette for our plot
palette( c("gray","pink") )

plot( res$log2FoldChange,  -log(res$padj), 
 col=res$sig+1, ylab="-Log(P-value)", xlab="Log2(FoldChange)")

# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.1), col="darkgray", lty=2)
```


```{r}
# Reset the color palette
palette("default")

plot( res$log2FoldChange,  -log(res$padj), 
 col=res$sig+1, ylab="-Log(P-value)", xlab="Log2(FoldChange)")

# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.1), col="darkgray", lty=2)
```


```{r}
# Setup our custom point color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

# padj is pvalue adjust, can choose what p value you would like to be plotted

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

#Volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)




```






SAVE OUR RESULTS

```{r}
write.csv(res, file = "expression_results.csv")

```





