---
title: "Cancer/Cancer Genomics"
author: "Justin"
date: "11/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

GenomicDataCommons R package

```{r, eval = FALSE}
# install/load packages

BiocManager::install( c("GenomicDataCommons", "TCGAbiolinks", "maftools") )

```

```{r}
library(GenomicDataCommons)
library(TCGAbiolinks)
library(maftools)
```

Check on GDC Status
```{r}
status()
```


```{r}
projects <- getGDCprojects()
head(projects)
```



Add data

```{r}
cases_by_project <- cases() %>%
  facet("project.project_id") %>%
  aggregations()
head(cases_by_project)
```


Make barplot
If you use the View(projects) function call you can see all the project names (such as Neuroblastoma, Pancreatic Adenocarcinoma, etc.) along with their project IDs (such as TARGET-NBL, TCGA-PAAD, etc.) and associated information.

Moving onto cases() we can use an example from the package associated publication to answer our first from question above (i.e. find the number of cases/patients across different projects within the GDC):

```{r}
x <- cases_by_project$project.project_id

# Make a custom color vector for our plot
colvec <- rep("lightblue", nrow(x))

# want to make the column that has "TCGA-PAAD" in the key col red, 
colvec[ (x$key == "TCGA-PAAD")] <- "red"

# Plot with 'log' for y axis and rotate labels with 'las'
#par(___)  

# Using log scale because there's a huge gap in the amount of data 
par(mar = c(5, 5, 5, 2.1))
barplot(x$doc_count, names.arg= x$key , log="y", col=colvec, las=2)

```


```{r}
rep
```




PART 2

```{r}

seqs <- read.fasta("lecture18_sequences.fa")


## Calculate positional identity scores
ide <- conserv(seqs$ali, method="identity")
mutant.sites <- which(ide < 1) 

## Exclude gap possitions from analysis
gaps <- gap.inspect(seqs)
mutant.sites <- mutant.sites[mutant.sites %in% gaps$f.inds]

mutant.sites

```

```{r}

# This will tell us what position the mutation is at because ide tells us which is less than 1 which means different
pos <-  which(ide < 1)[1]
pos

# we know that we want +/- 8 from the mutated sight position
(pos-8):(pos+8)
```



