---
title: "Class 7 R functions/practice"
author: "Justin"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**#R funxns Revisited**

Source functions from last class.

```{r}
# source() runs another file, allows u to pull something from another locxn or sumn; Pushing play button sources the HTML below and adds them to Global Environment

source("http://tinyurl.com/rescale-R")

```


```{r}
# Test funxn out

rescale(1:10)

# Test more; works with NA bc rescale default is na.rm = TRUE

rescale(c(1,10,5,NA,6))

```

```{r}
# Working with is.numeric and ! to flip logical

is.numeric(c(5:10))

# is.numeric only gives TRUE with numbers

is.numeric(c(5:10,"chocolate"))

# flip using ! to make it say TRUE with not numbers

!is.numeric(c(5:10,"chocolate"))



```


```{r}
# Test rescale2 with characters; input isn't numeric, rescale2 tells you it should be (author wrote that themselves)

# rescale2(c(5:10, "barry"))
```


**##Write Funxn to Find NAs**

1st make simple input where answer is known

```{r}

x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

# Test is.na

is.na(x)


# Using which() on is.na will give the positions where NA = TRUE

which(is.na(x))

which(is.na(y))


# Do more stuff

is.na(x)
is.na(y)

is.na(x) & is.na(y)

# use sum() to sum values; FALSEs(=0) and TRUEs (=1) are numbers

sum(is.na(x) & is.na(y))



```


**#Create Funxn
Make funxn using **sum()** and **is.na()** from before
```{r}
both_na <- function(x,y){
  sum( is.na(x) & is.na(y) )
}

```


Test the funxn we just created.
```{r}
both_na(x,y)
```

Add new variables from slides; and test funxn.
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

# Use both_na on x and y1

both_na( x, y1 )

# use both_na on x and y2

both_na( x, y2)




```


```{r}
x2 <- c( NA, NA )

both_na( x2, y2 )

```

Showing how stuff gets recycled or something
```{r}
plot(1:10, col = c("red", "blue", "green"))
```

Mess around w it
```{r}
x <- c( NA, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA, NA)

both_na(x,y3)
```

What happens is the first (shorter vector, will recycle itself to become as long as the @nd, longer vector)

```{r}
x3 <- c(1, NA, NA)
#what happens is
#x3<- c(1, NA, NA, 1,  NA, NA, 1)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
```



```{r}
length(x)
length(y3)
```

Make **both_na2** in conjunction with the **stop()** funxn to allow for an error message
```{r}

# Create both_na2

both_na2 <- function(x,y){
  if (length(x) != length(y) ) {
    stop("Inputs x and y should be the same length")
  }
}

# Test funxn

#both_na2(x, y3)


```
##Write grading funxn

```{r}
# Student 1 scores

a <- c(100, 100, 100, 100, 100, 100, 100, 90)

# Student 2 scores

b <- c(100, NA, 90, 90, 90, 90, 97, 80)

# using - gives everything BUT the value the original funxn gives
a[-which.min(a)]

b[-which.min(b)]

# Do mean of it
mean(a[ -which.min(a) ])

mean(b[ -which.min(b) ], na.rm = TRUE)



```
check how many NAs there are

```{r}

any(is.na(b))
```


**#Create the GODLIEST funxn EVER**

```{r}
a <- c(100, 100, 100, 100, 100, 100, 100, 90)
grade <- function(x){
  if(any(is.na(x))) {
    warning("Student missing HW")
    
  }
  mean(x[ -which.min(x) ], na.rm = TRUE)
  
}

grade(a)

grade(b)
```

Grade whole class
```{r}


class <- read.csv("https://tinyurl.com/gradeinput", row.names = 1)

# Use apply, 1 = row, 2 would be col; apply(x, margin, FUN)
apply(class, 1, grade)



```


datapasta Rcade

```{r}
# Rcade
# install.packages("devtools")
# devtools::install_github('RLesur/Rcade')

# USAGE
#  Rcade::games
#  Rcade::games$Pacman





```






















